---
title: "Multicollinearity Analysis"
author:
  - Katie Birchard
  - Ryan Homer
  - Andrea Lee
output:
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: true
bibliography: ../../doc/refs.bib
---

```{r libraries, message=FALSE}
library(tidyverse)
library(lubridate)
library(caret)
library(feather)
library(reshape2)
library(kableExtra)
library(car)
source(here::here("src/multicoll/support_functions.R"))
```

## Correlation Matrix

Get training data.

```{r data}
# See README for how to create feather files.
df <- read_feather(here::here("data/train.feather"))
```

Under the assumption that the data can be modelled linearly after observing the residual plot, we select the continuous numerical predictors, compute the correlation matrix and wrangle into a plottable dataframe [@corr].

```{r dataframe}
# Create dataframe for correlation matrix chart
corr_df <- df %>%
  select(total_volume,
         PLU_4046,
         PLU_4225,
         PLU_4770,
         total_bags,
         small_bags,
         large_bags,
         xlarge_bags) %>%
  cor %>%
  upper_tri_na %>%
  melt(na.rm = TRUE) %>%
  mutate(value = round(value, 2))
```

Correlation Matrix Chart

```{r chart}
ggplot(corr_df) +
  geom_tile(aes(Var1, Var2, fill = value)) +
  geom_text(aes(Var1, Var2, label = value), color = "black", size = 4) +
  scale_fill_distiller(palette = "GnBu", direction = 1) +
  labs(title = "Correlation Matrix",
       subtitle = "Avocado Data",
       fill = "Pearson's\nCorrelation") +
  coord_fixed() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.55, 0.7),
        legend.direction = "horizontal") +
  guides(fill = guide_colorbar(barwidth = 7,
                               barheight = 1,
                               title.position = "top",
                               title.hjust = 0.5))
```

Overall, there is fairly high collinarity between many of the predictors. This was expected, since they all deal with volume of avocados sold, be it by PLU code, bag type or total volume.

In particular, `total_bags` and `total_volume` were expected to be highly correlated to other predictors that were sub-quantities of these totals.

Due to the high correlation, including all these predictors in a prediction model would probably lead to overfitting.

## Multicollinearity

Create linear model and comput VIF scores from car [@car] package.

```{r mc 1}
lm(average_price ~ total_volume +
                   PLU_4046 +
                   PLU_4225 +
                   PLU_4770 +
                   total_bags +
                   small_bags +
                   large_bags +
                   xlarge_bags, data = df) %>%
  vif() %>%
  enframe() %>%
  rename(Predictor = name, "VIF Score" = value) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

This suggests extremely high collinearity for these variables in a linear model. We'll be careful about using these features. They are probably not very good predictors of the average avocado price.

## References
